---
title: "Data Summary"
author: "Matheus Hoffmann Fernandes Santos"
date: "April 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(sqldf)
```

## Collect

Create data frames from tables, which are comorbidity, demography, non-hematological and SNP(single nucleotide polymorphism)

```{r, echo=TRUE}
setwd("C:\\Users\\Hoffmann\\Desktop\\dados_dissert")

df_como <- read.csv("comorbidade.csv", sep = ";", stringsAsFactors = FALSE)
df_demo <- read.csv("dadosdemo.csv", sep = ";",stringsAsFactors = FALSE)
df_nhem <- read.csv("nhemato.csv", sep = ";", stringsAsFactors = FALSE)
df_snp <- read.csv("snpcorrect.csv", sep = ";",stringsAsFactors = FALSE)

cat("Table Summary: Observations and Variables\n")
cat(sprintf("Comorbidity ------------ %i | %i\n",nrow(df_como),ncol(df_como)))
cat(sprintf("Demographic ------------- %i | %i\n",nrow(df_demo),ncol(df_demo)))
cat(sprintf("Non-hemotological ------ %i | %i\n",nrow(df_nhem),ncol(df_nhem)))
cat(sprintf("SNP -------------------- %i | %i\n",nrow(df_snp),ncol(df_snp)))
```

## Missing Values

Showing percentage of missing values, by table.

### Comorbidity data

```{r, echo=TRUE}
#IMC has one value 999
count <- sqldf("SELECT COUNT(*) AS total FROM df_como")

queryBuilderMissing <- function(column, total, df){
 if(column == "IMC"){
      result <- sprintf("SELECT (1 - CAST(COUNT(%s) AS FLOAT)/%i) AS missing FROM %s WHERE %s < 60 OR %s <> '' OR %s <> NULL", #remove 999 value
                    column, total, df,column, column, column)
  } else {
      result <- sprintf("SELECT (1 - CAST(COUNT(%s) AS FLOAT)/%i) AS missing FROM %s WHERE %s <> '' OR %s <> NULL",
                    column, total, df, column, column)
  }

  #print(result)
  return(result)
}
getMissingByHeader <- function(){
  results <- c()
  percent <- "%"
  for(index in colnames(df_como)){
    missing_in_column <- sqldf(queryBuilderMissing(index, count$total, "df_como"))
    print(sprintf("%s missing values ratio >> %.2f%s", index, missing_in_column*100, percent))
    #results <- c(results, missing_in_column)
  }
  #return(results)
}

getMissingByHeader()

```

### Demographic data 

```{r, echo=TRUE}
count <- sqldf("SELECT COUNT(*) AS total FROM df_demo")
#One register filho = 99
queryBuilderMissing <- function(column, total, df){
 if(column == "filho"){
      result <- sprintf("SELECT (1 - CAST(COUNT(%s) AS FLOAT)/%i) AS missing FROM %s WHERE %s < 20 OR %s <> '' OR %s <> NULL", #remove 999 value
                    column, total, df,column, column, column)
  } else {
      result <- sprintf("SELECT (1 - CAST(COUNT(%s) AS FLOAT)/%i) AS missing FROM %s WHERE %s <> '' OR %s <> NULL",
                    column, total, df, column, column)
  }
  #print(result)
  return(result)
}
getMissingByHeader <- function(){
  results <- c()
  percent <- "%"
  for(index in colnames(df_demo)){
    missing_in_column <- sqldf(queryBuilderMissing(index, count$total, "df_demo"))
    print(sprintf("%s missing values ratio >> %.2f%s", index, missing_in_column*100, percent))
    #results <- c(results, missing_in_column)
  }
  #return(results)
}

getMissingByHeader()

```

### Single Nucleotide Polymorphism data

Old column names renamed to allow sqldf use, columnnames containing some special chars are mapped to '.'
Changes: '>' = bt, ' '='', '-'='' 
Obs: Numeric (first letter)headers are mapped to X{{Name}} 
```{r, echo=TRUE}
count <- sqldf("SELECT COUNT(*) AS total FROM df_snp")

queryBuilderMissing <- function(column, total, df){
  if(column == "N" || column == "Pront"){
          result <- sprintf("SELECT (1 - CAST(COUNT(%s) AS FLOAT)/%i) AS missing FROM %s WHERE %s<>'' OR %s<>NULL", column, total, df, column, column)
  } else {
      result <- sprintf("SELECT (1 - CAST(COUNT(%s) AS FLOAT)/%i) AS missing FROM %s WHERE %s='0' OR %s='1' OR %s='2'", column, total, df, column, column, column)
  }

  #print(result)
  return(result)
}
getMissingByHeader <- function(){
  #colnames old c("N", "Pront", "15631 G>T", "18053 A>G", "25505C>T", "6986 A>G", "C1236T", "G2677T/A", "C3435T", "A>G (ILE105VAL)", "02 (Ala114Val)", "11 G>A", "Ile655Val A>G", "A>G", "C>G")
  #results <- c()
  percent <- "%"
  for(index in colnames(df_snp)){
    missing_in_column <- sqldf(queryBuilderMissing(index, count$total, "df_snp"))
    print(sprintf("%s missing values ratio >> %.2f%s", index, missing_in_column*100, percent))
    #results <- c(results, missing_in_column)
  }
  #return(results)
}

getMissingByHeader()
```
### Non-hematological data

Check if percentage by reaction make more sense.
```{r, echo=TRUE}
count <- sqldf("SELECT COUNT(*) AS total FROM df_nhem")

queryBuilderMissing <- function(column, total, df){
  if(column == "Prontuario" || column == "Protocolo"){
          result <- sprintf("SELECT (1 - CAST(COUNT(%s) AS FLOAT)/%i) AS missing FROM %s WHERE %s<>'' OR %s<>NULL", column, total, df, column, column)
  } else {
      result <- sprintf("SELECT (1 - CAST(COUNT(%s) AS FLOAT)/%i) AS missing FROM %s WHERE %s='0' OR %s='1' OR %s='2' OR %s='3' OR %s='4'", column, total, df, column, column, column, column, column)
  }

  #print(result)
  return(result)
}
getMissingByHeader <- function(){
  percent <- "%"
  for(index in colnames(df_nhem)){
    missing_in_column <- sqldf(queryBuilderMissing(index, count$total, "df_nhem"))
    print(sprintf("%s missing values ratio >> %.2f%s", index, missing_in_column*100, percent))
    #results <- c(results, missing_in_column)
  }
  #return(results)
}

getMissingByHeader()
```

